---
title: "Examine Life Expectancy"
author: "Rachel Regberg"
date: "`r Sys.Date()`"
output: html_document
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Here, I re-examine the analysis of Oeppen and Vaupel (2002).


## Life Expectancy Analysis

We will first clean the data.

1. Load the `gapminder_full` data fram into R. 
```{r warning=FALSE}
suppressPackageStartupMessages(library(tidyverse))
library(gapminder)
data("gapminder_unfiltered")
```

2. Calculate the maximum life expectancy each year and the corresponding country that had that maximum life expectancy. Hint: Use `group_by()` and `summarize()`.
```{r}
gapminder_unfiltered %>%
  group_by(year, country) %>%
  summarize(maxLifeExp=max(lifeExp)) %>%
  slice(which.max(maxLifeExp)) -> MaxLifeExp
MaxLifeExp
```

Alternate/Correct solution:
```{r}
gapminder_unfiltered %>%
  group_by(year) %>%
  filter(lifeExp==max(lifeExp)) %>%
  select(country, lifeExp, year) %>%
  arrange(year) ->
  LifeExpectancy
LifeExpectancy
```

We will now recreate the plot from Oeppen and Vaupel (2002).

My answer:
```{r}
gapminder_unfiltered %>%
  group_by(year, country) %>%
  summarize(maxLifeExp=max(lifeExp)) %>%
  slice(which.max(maxLifeExp)) %>%
  ggplot(aes(x=year, y=maxLifeExp)) +
  geom_point(mapping=aes(fill=country), size=3, pch=21) +
  geom_smooth(se=FALSE, method=lm, color="blue") +
  scale_color_discrete(name="Country") +
  xlab("Year") +
  ylab("Maximum Life Expectancy") +
  theme_bw()
```

In class answer:
```{r}
LifeExpectancy %>%
  ggplot(aes(x=year, y=lifeExp, color=country)) +
  geom_smooth(se=FALSE, method="lm", color="blue") +
  geom_point(pch=1)
```
