---
title: "Examine Life Expectancy"
author: "Rachel Regberg"
date: "`r Sys.Date()`"
output: html_document
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Here, I re-examine the analysis of Oeppen and Vaupel (2002).


## Life Expectancy Analysis

We will first clean the data.

1. Load the `gapminder_full` data fram into R. 
```{r warning=FALSE}
suppressPackageStartupMessages(library(tidyverse))
library(gapminder)
data("gapminder_unfiltered")
```

2. Calculate the maximum life expectancy each year and the corresponding country that had that maximum life expectancy. Hint: Use `group_by()` and `summarize()`.
```{r}
gapminder_unfiltered %>%
  group_by(year, country) %>%
  summarize(maxLifeExp=max(lifeExp)) %>%
  slice(which.max(maxLifeExp)) -> MaxLifeExp
MaxLifeExp
```

Alternate/Correct solution:
```{r}
gapminder_unfiltered %>%
  group_by(year) %>%
  filter(lifeExp==max(lifeExp)) %>%
  select(country, lifeExp, year) %>%
  arrange(year) ->
  LifeExpectancy
LifeExpectancy
```

We will now recreate the plot from Oeppen and Vaupel (2002).
